Supplemental Information A: Full Annotated Analysis and Reproducible Code

### 2023-07-28 Supplemental-Material-A-Annotated-Analysis.Rmd
adj SI: A

---
title: "Supplemental Information A: Full Annotated Analysis and Reproducible Code"
subtitle: "Supplemental Material for 'Migration Trajectories [working title]'"
author:
- Jannis Kreienkamp^1,3^
- Kai Epstude^1,3^
- Laura F. Bringmann^1,3^
- Maximilian Agostini^1,3^
- Peter de Jonge^1,3^
- Rei Monden^2,3^
- ^1^University of Groningen, Department of Psychology
- ^2^Hiroshima University, Graduate School of Advanced Science and Engineering
- ^3^Author order TBD [currently in first name alphabetical order]
- "Author Information:"
- "Correspondence concerning this article should be addressed to Jannis Kreienkamp, Department of Psychology, University of Groningen, Grote Kruisstraat 2/1, 9712 TS Groningen (The Netherlands).  E-mail: j.kreienkamp@rug.nl"
- 'The main manuscript is available at <a href="https://www.doi.org/ToBePublished" target="_blank">doi.org/ToBePublished</a>'
- 'The data repository for this manuscript is available at <a href="https://osf.io/TBA" target="_blank">osf.io/TBA</a>'
- 'The GitHub repository for this manuscript is available at <a href="https://github.com/maskedForPeerReview" target="_blank">github.com/maskedForPeerReview</a>'
date: "Last updated: `r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::html_document2:
    fig_caption: yes
    md_extensions: +footnotes
    code_folding: hide
    mathjax: default
    theme: yeti
    toc: yes
    toc_float: yes
    number_sections: false
    css: style.css
    includes:
      in_header: "_includes/head-custom-rmd.html" 
editor_options:
  chunk_output_type: console
bibliography: references.bib
csl: apa.csl
header-includes:
   - \usepackage{amsmath, nccmath}
---

```{=html}
<style type="text/css">
.main-container {
  max-width: 1300px;
  margin-left: auto;
  margin-right: auto;
}
.table {
  margin-left:auto; 
  margin-right:auto;
}
</style>
```

@@

# Global Chunk Options
knitr::opts_chunk$set(
  fig.width = 12,
  fig.height = 8,
  fig.path = "Figures/",
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

@@

The data were collected using two different survey tools. For the study with sojourners (Study 1: worker) we used the survey platform Qualtrics XM, whereas the studies with international students (Study 2: student), and the international medical professionals (Study 3: medical) were conducted using the survey framework FormR. This means that the datasets had inconsistent file formats and naming conventions. For the Qualtrics study we pre-processed some variables to ease the import process (for the syntax files see the SPS files in '*data/S1_Workers/processed/cleaned*' and for the raw data files see '*data/S1_Workers/raw*'). For the two other studies, we import the raw csv files from their respective folders.

@@

For the student sample data was, similarly, collected in three separate surveys: (1) the pre-measurement, (2) the daily survey sent out at lunch and dinner time, and (3) a post-measurement. We combine the three individual surveys into one large dataframe and drop superfluous variables that are not relevant to the analyses relevant here. We exclude our own test data as well as one participant who entered the study twice (but gave different responses during the pre-measurement). We also reformat missing values and format core ID variables.\
<i class="fas fa-info-circle"></i> *Note:* All data preparation steps are saved in the '*dtS2*' list.

@@

# Prepare dfs for Cleaning
df.pre <- dtS2$raw.pre %>%
  mutate_all(na_if, "") %>%
  mutate_all(na_if, "NA") %>%
  filter(!is.na(ended)) %>% # remove all who did not finish
  filter(!email %in% .$email[duplicated(.$email)]) %>% # remove all who did the pre questionnaire multiple times (b/c inconsistent ratings scales)
  filter(!ResponseId %in% ownIDs) %>% # remove our own test
  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
names(df.pre) <- paste(names(df.pre), "pre", sep = ".")

df.post <- dtS2$raw.post %>%
  mutate_all(na_if, "") %>%
  mutate_all(na_if, "NA") %>%
  filter(!is.na(ResponseId)) %>% # remove own test runs
  filter(!ResponseId %in% ownIDs) %>% # remove our own test
  filter(ResponseId %in% df.pre$ResponseId) %>% # remove anyone who wasn't in the pre
  filter(!is.na(ended)) %>% # remove all who never finished
  filter(!ResponseId %in% .$ResponseId[duplicated(.$ResponseId)]) %>% # remove all duplicate ResponseIds
  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
names(df.post) <- paste(names(df.post), "post", sep = ".")

df.daily <- dtS2$raw.daily %>%
  mutate_all(na_if, "") %>%
  mutate_all(na_if, "NA") %>%
  filter(!ResponseId %in% ownIDs) %>% # remove our own test
  filter(ResponseId %in% df.pre$ResponseId) %>% # remove anyone who wasn't in the pre
  filter(!is.na(ended)) %>% # remove all who never finished
  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)

@@

# Prepare dfs for Cleaning
df.pre <- dtS3$raw.pre %>%
  mutate_all(na_if, "") %>%
  mutate_all(na_if, "NA") %>%
  filter(!is.na(ended)) %>% # remove all who did not finish
  filter(!ResponseId %in% ownIDs) %>% # remove our own test
  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
names(df.pre) <- paste(names(df.pre), "pre", sep = ".")

df.post <- dtS3$raw.post %>%
  mutate_all(na_if, "") %>%
  mutate_all(na_if, "NA") %>% 
  filter(!is.na(ResponseId)) %>% # remove own test runs
  filter(!ResponseId %in% ownIDs) %>% # remove our own test
  filter(ResponseId %in% df.pre$ResponseId) %>% # remove anyone who wasn't in the pre
  #filter(!is.na(ended)) %>% # remove all who never finished [disabled because only relevant if data is missing]
  filter(!ResponseId %in% .$ResponseId[duplicated(.$ResponseId)]) %>% # remove all duplicate ResponseIds
  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
names(df.post) <- paste(names(df.post), "post", sep = ".")

df.daily <- dtS3$raw.daily %>%
  mutate_all(na_if, "") %>%
  mutate_all(na_if, "NA") %>%
  filter(!ResponseId %in% ownIDs) %>% # remove our own test
  filter(ResponseId %in% df.pre$ResponseId) %>% # remove anyone who wasn't in the pre
  #filter(!is.na(ended)) %>% # remove all who never finished [disabled because only relevant if data is missing]
  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)

@@

1. Variable selection
2. Feature extraction
3. (Feature selection)
4. Clustering
5. Visualization, interpretation, and validation

@@

We begin by selecting the variables, which we would like to include in the clustering procedure. For the clustering of time-series characteristics we focus on variables that are specific to the ESM measurements (i.e., variables collected during the various daily dairy pings). This means, that variables from the pre- and post-test are excluded from the clustering procedure. While the variables can in theory be added as additional features, they are not the focus of this manuscript.

@@

apply(dtS1Red, 2, sum(is.na(.))/nrow(.))

your_dataframe %>%
  group_by(Your_Participantid) %>%
  summarise(across(everything(), funs(100*mean(is.na(.)))))

@@

var <- get_pca_var(res.pca)
corrplot::corrplot(var$contrib[,1:10], is.corr=FALSE) 
fviz_contrib(res.pca, choice = "var", axes = 1, top = 50)
fviz_contrib(res.pca, choice = "var", axes = 2, top = 50)

res.desc <- FactoMineR::dimdesc(res.pca, axes = c(1,2), proba = 0.05)
# Description of dimension 1
res.desc$Dim.1

>>>>

---
title: "Supplemental Information A: Full Annotated Analysis and Reproducible Code"
subtitle: "Supplemental Material for 'Migration Trajectories [working title]'"
author:
- Jannis Kreienkamp^1,3^
- Kai Epstude^1,3^
- Laura F. Bringmann^1,3^
- Maximilian Agostini^1,3^
- Peter de Jonge^1,3^
- Rei Monden^2,3^
- ^1^University of Groningen, Department of Psychology
- ^2^Hiroshima University, Graduate School of Advanced Science and Engineering
- ^3^Author order TBD [currently in first name alphabetical order]
- "Author Information:"
- "Correspondence concerning this article should be addressed to Jannis Kreienkamp, Department of Psychology, University of Groningen, Grote Kruisstraat 2/1, 9712 TS Groningen (The Netherlands).  E-mail: j.kreienkamp@rug.nl"
- 'The main manuscript is available at <a href="https://www.doi.org/ToBePublished" target="_blank">doi.org/ToBePublished</a>'
- 'The data repository for this manuscript is available at <a href="https://osf.io/TBA" target="_blank">osf.io/TBA</a>'
- 'The GitHub repository for this manuscript is available at <a href="https://github.com/maskedForPeerReview" target="_blank">github.com/maskedForPeerReview</a>'
date: "Last updated: `r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::html_document2:
    highlight: pygments
    fig_caption: yes
    md_extensions: +footnotes
    code_folding: hide
    mathjax: default
    theme: yeti
    toc: yes
    toc_float: yes
    number_sections: false
    css: style.css
    includes:
      in_header: "_includes/head-custom-rmd.html" 
editor_options:
  chunk_output_type: console
bibliography: referencesZotero.bib
csl: apa.csl
header-includes:
   - \usepackage{amsmath, nccmath}
---

```{=html}
<style type="text/css">
.main-container {
  max-width: 1300px;
  margin-left: auto;
  margin-right: auto;
}
.table {
  margin-left:auto; 
  margin-right:auto;
}

body
  { counter-reset: source-line 0; }
pre.numberSource code
  { counter-reset: none; }

</style>
```

@@

# Global Chunk Options
knitr::opts_chunk$set(
  fig.width = 12,
  fig.height = 8,
  fig.path = "Figures/",
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  class.source = "numberLines lineAnchors"
)
```

@@

The data were collected using two different survey tools. For the study with sojourners (Study 1: worker) we used the survey platform Qualtrics XM, whereas the studies with international students (Study 2: student), and the international medical professionals (Study 3: medical) were conducted using the survey framework FormR. This means that the datasets had inconsistent file formats and naming conventions. For the Qualtrics study we pre-processed some variables to ease the import process (for the syntax files see the SPSS files in '*data/S1_Workers/processed/cleaned*' and for the raw data files see '*data/S1_Workers/raw*'). For the two other studies, we import the raw csv files from their respective folders.

@@

For the student sample data were, similarly, collected in three separate surveys: (1) the pre-measurement, (2) the daily survey sent out at lunch and dinner time, and (3) a post-measurement. We combine the three individual surveys into one large dataframe and drop superfluous variables that are not relevant to the analyses relevant here. We exclude our own test data as well as one participant who entered the study twice (but gave different responses during the pre-measurement). We also reformat missing values and format core ID variables.\
<i class="fas fa-info-circle"></i> *Note:* All data preparation steps are saved in the '*dtS2*' list.

@@

# Prepare dfs for Cleaning
df.pre <- dtS2$raw.pre %>%
  mutate(across(where(is.character), na_if, "")) %>%
  mutate(across(where(is.character), na_if, "NA")) %>%
  filter(!is.na(ended)) %>% # remove all who did not finish
  filter(!email %in% .$email[duplicated(.$email)]) %>% # remove all who did the pre questionnaire multiple times (b/c inconsistent ratings scales)
  filter(!ResponseId %in% ownIDs) %>% # remove our own test
  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
names(df.pre) <- paste(names(df.pre), "pre", sep = ".")

df.post <- dtS2$raw.post %>%
  mutate(across(where(is.character), na_if, "")) %>%
  mutate(across(where(is.character), na_if, "NA")) %>%
  filter(!is.na(ResponseId)) %>% # remove own test runs
  filter(!ResponseId %in% ownIDs) %>% # remove our own test
  filter(ResponseId %in% df.pre$ResponseId) %>% # remove anyone who wasn't in the pre
  filter(!is.na(ended)) %>% # remove all who never finished
  filter(!ResponseId %in% .$ResponseId[duplicated(.$ResponseId)]) %>% # remove all duplicate ResponseIds
  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
names(df.post) <- paste(names(df.post), "post", sep = ".")

df.daily <- dtS2$raw.daily %>%
  mutate(across(where(is.character), na_if, "")) %>%
  mutate(across(where(is.character), na_if, "NA")) %>%
  filter(!ResponseId %in% ownIDs) %>% # remove our own test
  filter(ResponseId %in% df.pre$ResponseId) %>% # remove anyone who wasn't in the pre
  filter(!is.na(ended)) %>% # remove all who never finished
  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)

@@

# Prepare dfs for Cleaning
df.pre <- dtS3$raw.pre %>%
  mutate(across(where(is.character), na_if, "")) %>%
  mutate(across(where(is.character), na_if, "NA")) %>%
  filter(!is.na(ended)) %>% # remove all who did not finish
  filter(!ResponseId %in% ownIDs) %>% # remove our own test
  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
names(df.pre) <- paste(names(df.pre), "pre", sep = ".")

df.post <- dtS3$raw.post %>%
  mutate(across(where(is.character), na_if, "")) %>%
  mutate(across(where(is.character), na_if, "NA")) %>%
  filter(!is.na(ResponseId)) %>% # remove own test runs
  filter(!ResponseId %in% ownIDs) %>% # remove our own test
  filter(ResponseId %in% df.pre$ResponseId) %>% # remove anyone who wasn't in the pre
  #filter(!is.na(ended)) %>% # remove all who never finished [disabled because only relevant if data is missing]
  filter(!ResponseId %in% .$ResponseId[duplicated(.$ResponseId)]) %>% # remove all duplicate ResponseIds
  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
names(df.post) <- paste(names(df.post), "post", sep = ".")

df.daily <- dtS3$raw.daily %>%
  mutate(across(where(is.character), na_if, "")) %>%
  mutate(across(where(is.character), na_if, "NA")) %>%
  filter(!ResponseId %in% ownIDs) %>% # remove our own test
  filter(ResponseId %in% df.pre$ResponseId) %>% # remove anyone who wasn't in the pre
  #filter(!is.na(ended)) %>% # remove all who never finished [disabled because only relevant if data is missing]
  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)

@@

1. Variable selection
2. Feature extraction
3. Feature reduction
4. Clustering
5. Visualization, interpretation, and validation

@@

We included 12 variables that were available across all three studies and captured information about the participant's interactions, as well as cognitive-, emotional-, and motivational self in relationship with the majority group. We chose these two aspects in particular because (1) the interaction-specific information exemplified the structural missingness issue of modern ESM data and (2) the motivational, emotional, and cognitive experience offered a diverse conceptualization of migration experience (beyond behavioral measurements) that is becoming more common in the literature [@Kreienkamp2022d].

FILTER BY main

@@

# apply(dtS1Red, 2, sum(is.na(.))/nrow(.))
# 
# your_dataframe %>%
#   group_by(Your_Participantid) %>%
#   summarise(across(everything(), funs(100*mean(is.na(.)))))

@@

# var <- get_pca_var(res.pca)
# corrplot::corrplot(var$contrib[,1:10], is.corr=FALSE) 
fviz_contrib(res.pca, choice = "var", axes = 1, top = 50)
fviz_contrib(res.pca, choice = "var", axes = 2, top = 50)

res.desc <-
  FactoMineR::dimdesc(
    FactoMineR::PCA(scaled_data, scale.unit = FALSE, graph = FALSE),
    axes = c(1, 2),
    proba = 0.05
  )
# Description of dimension 1
# res.desc$Dim.1
